---
title: "5.Hierarchical clustering & subtype discovery"
author: "yueh-ting"
date: "2020/2/15"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
##1531 gene from step4
#use the file from step4,"newpass_gene.csv", and make the dataframe
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
df <- read.csv("newpass_gene.csv", row.names=1)
require(data.table)
rdf <- transpose(df)
```

#do the cluster
use euclidean & Ward method to cluster, here I use the most common way.
and I got 57 for one cluster, 77 for another cluster
```{r}
D <- dist(rdf,method = 'euclidean')
hclu_res <- hclust(D,method = "ward.D2")
clusterCut <- cutree(hclu_res, 2)
sum(clusterCut == 1)
sum(clusterCut == 2)
```

## use the proj_metadata.csv to pair the C3 data
include the other file to get the name.
```{r ,include=FALSE}
proj_metadata <- read.csv("proj_metadata.csv")
match_pair <- proj_metadata$geo_accession
match <- c()
for (i in 1:dim(df)[2]){
  match[i] <- substr(colnames(df)[i],1,9)
}
match <- as.factor(match)
match == match_pair
```

# column specific color and do the heat map

```{r}
color = ifelse(proj_metadata$cit.coloncancermolecularsubtype == 'C3','red','blue')
heatmap(as.matrix(df),ColSideColors = color)
```

# seperate G1&G2 to Group1 and Group2 for later to do the t test
```{r}
G1 <- df[,which(clusterCut == 1)]
G2 <- df[,which(clusterCut == 2)]
```

# Welth t test (two group variance are not equal and unknown)
#Here I add the 5th column to count the adjust-pvalue that less than 0.05
#use a for loop to go through each value

```{r}
Gtable <- matrix(0,ncol=5,nrow=dim(df)[1])
for (i in 1:dim(df)[1]){
  res <- t.test(G1[i,],G2[i,])
  p_adj <- p.adjust(res$p.value,method = 'fdr',n = dim(df)[1])
  Bool <- ifelse(p_adj < 0.05,1,0)
  Gtable[i,] <- c(row.names(df)[i],as.numeric(res$statistic),res$p.value,p_adj,Bool)}

```

#store the file that pass the 0.05, there are 1012 genes
```{r}
Final <- Gtable[which(Gtable[,5] == 1),]
write.csv(Final,"5-3_pass.csv")
#store the file that have all the gene
write.csv(Gtable[,],"5-4_All_list.csv") 
rankfinal<- Final[order(as.numeric(Final[,4])), ]
write.csv(rankfinal, "rank_adjustpvalue.csv")
```{r}
